 \documentclass[
   titlepage=no]{scrartcl} %cuſtom titlepage: ſtill to come

 \usepackage{
   xltxtra, %xetex enhancements and fixes
   polyglossia, %hyphenation
   relsize, %\larger and \smaller text ſize
   verbatim, %quotes
   xcolor, %Driver-independent color extenſions
   xspace, %guesses space after commands
   csquotes %language ſpecific quotation marks for BibLaTeX
   }

 \usepackage[style=authoryear]{biblatex}%for the bibliography
 \bibliography{frunge-bibliography}%our database file

 \usepackage[
   breaklinks,
   colorlinks=true,
   pdfborder=0 0 0,
   pdftitle={Frunge Font Documentation},
   linkcolor=blue,
   urlcolor=blue,
   xetex]{hyperref}

%% font settings
 \setmainfont{Linux Libertine} %without O (this is a feature/dirty fix)
 \setsansfont[Scale=MatchLowercase]{Linux Biolinum O}
 
%% polyglossia settings
 \setmainlanguage{english}
 \frenchspacing
 \setotherlanguage[spelling=new, latesthyphen=true]{german}
 % pleaſe add babelshorthands=true if TeXLive 2009 is realeaſed

%% text commands
 \newcommand*\team{Dennis Heidſiek\and Chriſtian Kluge\and Martin Roppelt\and Arno 
Trautmann}

 \newcommand*\smiley{{\larger[3]☺}}
 \newcommand*\frowney{{\larger[3]☹}}
 \newcommand*\winkey{;)}

 \newcommand*\ipa[1]{\mbox{#1}} %another font would be nice
 \newcommand*\pkg[1]{\textsf{#1}}

  \newcommand\frungepangram{\textgerman{\IfFileExists{frunge-pangram.tex}{\input{frunge-pangram.tex}}{Pangramm-Platzhalter}}}


 %\usepackage{syntonly}
 %\syntaxonly

 %\includeonly{}


 \subject{The Documentation of} \title{The Frunge Font Family}
 \author{\team}

 \addtokomafont{section}{\LARGE}




 \begin{document}

 \begin{comment}
 Engliſh long s rules:
 End of word: s / Elſe: ſ
 with f: sf/fs
 with daſh at line end: ſ-
 with apostrophe: s’
 Italic writing: ſſ -> ſs
 17th century: ſb ſk -> sb sk
 \end{comment}


 \maketitle
 \thispagestyle{empty}
 \begin{abstract}
 \large Copyright 2009 Dennis Heidſiek 
(\href{mailto:heidsiekb@aol.com}{heidsiekb@aol.com}), Martin Roppelt 
\break (\href{mailto:m.p.roppelt@web.de}{m.p.roppelt@web.de}) and Arno 
Trautmann (\href{mailto:arno.trautmann@gmx.de}{arno.trautmann@gmx.de})

 Permiſſion is granted to copy, diſtribute and\,/\,or modify this 
document under the terms of the GNU Free Documentation Licenſe, Version 
1.3 or any later verſion publiſhed by the Free Software Foundation; with 
no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts. A 
copy of the licenſe is included in the ſection entitled 
“\hyperlink{fdl}{GNU Free Documentation Licenſe}”.

This work is licenſed under the Creative Commons Attribution-ShareAlike 
3.0 Unported Licenſe. To view a copy of this licenſe, viſit 
\href{http://creativecommons.org/licenses/by-sa/3.0/legalcode}{this 
page}; or, (b) ſend a letter to Creative Commons, 171 
2\textsuperscript{nd} Street, Suite 300, San Franciſco, California, 
94105, USA.

%If we plan to include (non-trivial?) ſource code:
 \begin{comment}
 This document is free ſoftware: you can rediſtribute it and\,/\,or 
modify it under the terms of the GNU General Public Licenſe as publiſhed 
by the Free Software Foundation, either verſion 3 of the Licenſe, or (at 
your option) any later verſion.

This document is diſtributed in the hope that it will be uſeful, but 
WITHOUT ANY WARRANTY; without even the implied warranty of 
MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General 
Public Licenſe for more details.

You ſhould have received a copy of the GNU General Public Licenſe along 
with this document. If not, see: http://www.gnu.org/licenses/gpl.html
 \end{comment}

 \end{abstract}
 \clearpage
 \tableofcontents
 \clearpage

 \addsec{Frunge Fonts Project}
 We at the Frunge Fonts Project want to create free, high-quality and 
Unicode complete ſubſtitutes to the proprietary blackletter fonts. 
Frunge is an abbreviation of the German “FRei UNd GEbrochen” (= free and 
blackletter).

Therefore it is pronounced \ipa{[fʁʊŋə]}. But as we tend to pronounce 
the Gentium SIL not as \ipa{[dʒɛntɪʊm]}, but rather as \ipa{[gɛntɪʊm]} 
or even \ipa{[gɛntsɪʊm]}, you may feel free to pronounce it as 
\ipa{[frʌndʒ]}, or as you like it.

We alſo found out, that ſome online dictionaries define frunge as “neck 
maſſage”. This is what it gives us when we finiſhed a font \smiley


\addsec{We Care About Your Freedom}
 Blackletter fonts are uſually very expenſive. Even if not, you may not 
ſtudy, change or even rediſtribute modified verſions \frowney

Becauſe we didn’t want to wait until thoſe publiſhers relicenſe their 
fonts as free ſoftware, we ſtarted this project. A little bit preſſure 
can help them, as well \winkey

Our fonts are free ſoftware, i.\,e. you may
 \begin{itemize}
  \item uſe
  \item ſtudy
  \item modify\,/\,improve and
  \item rediſtribute them.
 \end{itemize}

Uſually you can get them without fee.

We want to enſure that our fonts cannot be made proprietary again, ſo we 
choſe to licenſe them under a 
\href{http://en.wikipedia.org/wiki/Copyleft}{copyleft} licenſe:

To avoid licenſe proliferation and to enſure greateſt font combining 
poſſibilities, our fonts are licenſed under GPL, OFL and CC-BY-SA?.
 %If you feel that this is not enough, pleaſe 
\hyperlink{contact}{contact} us.

If you want to rediſtribute our fonts (probably changed), have a look at 
the full licenſe text(s) to make ſure, that you meet the conditions!

To maximize your freedom, we added exceptions to our licenſes to avoid 
ſome reſtrictions of the vanilla licenſes:

 GPL:
 \begin{itemize}
  \item you may embed this font

 OFL:
  \item you may uſe our font name and
  \item you may ſell our fonts unbundled with other ſoftware
 \end{itemize}


 \addsec{Feature Goals}
 \begin{itemize}
  \item \XeTeX\ Math Fonts
  %\item Language specific typography
  %\item r rotunda
  \item vowels with an “e” as umlaut mark
  \item no overwriting of existing Unicode codepoints
  %\item \href{http://www.e-welt.net/bfds_2003/bund/normsondz/Normblatt 
%UNZ 1 - Antiqua.pdf}{UNZ1} compatibility
 \end{itemize}


 \addsec{Font Faces}
 \addsec*{Frunge Mixtura}

This is a draft and learning example. Its outlines are drawn as if their 
ſtrokes were written with a thick feather, hold by 45 degrees (is to be 
decreased) only, in the lower and upper right direction.

It can be altered through options heavily (not implemented yet).


 \addsec{\texorpdfstring{\XeTeX-Usage}{XeTeX-Usage}}
 The uſage of the fonts with \XeTeX\ is quite eaſy. If you know how to 
uſe \TeX, just load the package 
\href{http://www.ctan.org/tex-archive/macros/xetex/latex/xltxtra/}{\pkg{xltxtra}} 
(which loads the \pkg{fontspec}-package, which does moſt of the work). 
Load the font ſaying \verb|\setmainfont{Frunge Font}|, where 
\verb|Frunge Font| is the name of the font as it is known to your 
operating ſyſtem. This will change the main font of your document to the 
given font. Say \verb|\fontspec{Frunge Font}| to change to the font 
locally.

The uſage of advanced OpenType features will be described in this place 
as ſoon as they are implemented in the fonts.
 %smcp
 %salt


 \addsec{Internal Implementation}
 Accented\,/\,Modified letters are diſſolved via a ccmp-table and 
reſurrected by abvm/blwm/mkmk/mark-tables. Therefore your font rendering 
engine ſhould fully ſupport OpenType (otherwiſe no glyphs are ſhown for 
thoſe). …


 \addsec{Typing of Long s}

%Xmodmap …

Neo—ergonomically optimized keyboard layout for German (and Engliſh); 
ſets long s on firſt level and makes blackletter typing eaſy. 
\href{http://wiki.neo-layout.org/wiki/Neo\%20unter\%20Linux\%20einrichten/Lang-s-Tastatur}{Inſtallation 
inſtructions for GNU/Linux [German]}


 \addsec{Testing of Fonts}
 You can teſt the Frunge fonts in this document with the command 
\verb|\frungepangram| wich will input the file \verb|frunge-pangram|:
 \begin{quote}
  \frungepangram
 \end{quote}
 This (non-ſenſical) text has all important glyphs to ſhow how a 
blackletter font would look like. The text may change from time to time 
as we might find better words \winkey


 \addsec{Similar Projects}
 \begin{itemize}
  \item 
\href{http://fontforge.ſourceforge.net/sfds/toyfonts.html}{FontForge’s 
Toy Fonts}
  \item 
\href{http://www.ctan.org/tex-archive/fonts/gothic/yfrak/}{Yannis 
Haralambous’s Blackletter Metafonts}
  \item \href{http://www.peter-wiegel.de/Fonts.html}{Peter Wiegel’s 
Blackletter and Kurrent fonts}
 \end{itemize}


 \addsec{Contact}
 If you want to contribute ſome patches or feature or bug requeſts, 
viſit our \href{http://ſv.nongnu.org/p/frunge}{project ſite}\\
 \hypertarget{contact}{You can write us email, too: 
\href{mailto:frunge-external@nongnu.org}{frunge-external@nongnu.org}}.\\ 
If you want to contribute\,/\,diſcuſs on a regular baſis, ſubſcribe to 
our 
\href{http://liſts.nongnu.org/mailman/listinfo/frunge-internal}{mailing 
liſt}.


 \printbibliography


 \include{fdl}


 \end{document}
